Token-level author diarization using clustering of stylistic contexts

### 2017-05-31 paper/paper.tex
Initial version of chapter 4.

\paragraph{Basic feature extraction.} 
We defined the stylistic context of a token $t_i$ as a set of tokens $\{t_k\}_{k=i-c}^{i-1} \cup \{t_k\}_{k=i+1}^{i+c}$ where $c$ is a context size. From each context we extracted the most used stylometric features which we found in previous work.
\begin{itemize}
	\item \textit{Character tri-grams}. Frequencies of \textit{n}-grams on character level have been very useful in quantifying the writing style \citep{stamatatos-2009b}. They are able to capture lexical and contextual information, use of punctuation and errors which can be an author's  "fingerprint". This feature is also tolerant to noise. Based on work by \citet{stamatatos-2009a} and \citet{rahman-2015}, we choose $n=3$. Maximal dimension of this feature vector was set to 200.
	\item \textit{Stop words}. According to \citet{stamatatos-2009a}, sometimes also called \textit{function words}, these are the most common used topic-independent words in text, such as articles, prepositions, pronouns and others. They are used unconciously and found to be one of the most discriminative features in authorship attribution since they represent pure stylistic choices of authors' {burrows-1987,argamon-2005}. We used frequencies of 156 english stop words available in \texttt{nltk}.
	\item \textit{Special characters}. We used counts of all character sequences which satisfied a regular expression defined as [REGEX]. Although character n-grams can catch the use of those character sequences, we wanted to have a distinct feature for that purpose. \citet{koppel-2009} mentioned that authors can have different punctuation habbits.
	\item \textit{POS tag counts}. This is syntactic feature which \citet{koppel-2009} and \citet{stamatatos-2009a} also identified as a discriminative one in authorship analysis and it was used by \citet{kuznetsov-2016}. We used all 12 tags from the universal tagset.
	\item \textit{Average token length}. Used by \citet{kuznetsov-2016}, \citet{sittar-2016},  \citet{brooke-2012} and \citet{stein-2011}. \citet{koppel-2009} characterized this feature as a complexity measure.
	\item \textit{Bag of Words}. Bag of words text representation more captures content, rather than style \citep{stamatatos-2009a}. We included this feature because it boosted performance of our initial testing, even without words been previously stemmed or lemmatized (stop words were excluded). Vocabulary had a maximum of 100 words.
	\item \textit{Type-token ratio}. We wanted to use that feature to measure the vocabulary richness in a token's context, but after we finished with performance evaluations we realized that there was a bug in implementation and that feature did not contribute to overal results. The feature should be calculated as the ratio of vocabulary size and total number of tokens of the text \citep{stamatatos-2009a}.
\end{itemize}

Differences between the 2 model variants....

basic features

@@

scaling

@@

clustering algorithms

@@

1
baselines
repo link
Mention cofidences.

>>>>

\paragraph{Basic feature extraction.} 
We defined the stylistic context of a token $t_i$ as a set of tokens $\{t_k\}_{k=i-cs}^{i-1} \cup \{t_k\}_{k=i+1}^{i+cs}$ where $cs$ is a context size. From each context we extracted the most used stylometric features which we found in previous work.
\begin{itemize}
	\item \textit{Character tri-grams}. Frequencies of \textit{n}-grams on character level have been very useful in quantifying the writing style \citep{stamatatos-2009b}. They are able to capture lexical and contextual information, use of punctuation and errors which can be an author's  "fingerprint". This feature is also tolerant to noise. Based on work by \citet{stamatatos-2009a} and \citet{rahman-2015}, we choose $n=3$. Maximal dimension of this feature vector was set to 200.
	\item \textit{Stop words}. According to \citet{stamatatos-2009a}, sometimes also called \textit{function words}, these are the most common used topic-independent words in text, such as articles, prepositions, pronouns and others. They are used unconciously and found to be one of the most discriminative features in authorship attribution since they represent pure stylistic choices of authors' \citep{burrows-1987,argamon-2005}. We used frequencies of 156 english stop words available in \texttt{NLTK}.
	\item \textit{Special characters}. We used counts of all character sequences which satisfied a regular expression defined as \verb/[^\\w]{1,4}/. Although character $n$-grams can catch the use of those character sequences, we wanted to have a distinct feature for that purpose since \citet{koppel-2009} mentioned that authors can have different punctuation habbits.
	\item \textit{POS tag counts}. This is syntactic feature which \citet{koppel-2009} and \citet{stamatatos-2009a} also identified as a discriminative one in authorship analysis and it was used by \citet{kuznetsov-2016}. We used all 12 tags from the universal tagset.
	\item \textit{Average token length}. Used by \citet{kuznetsov-2016}, \citet{sittar-2016},  \citet{brooke-2012} and \citet{stein-2011}. \citet{koppel-2009} characterized this feature as a complexity measure.
	\item \textit{Bag of Words}. Bag of words text representation more captures content, rather than style \citep{stamatatos-2009a}. We included this feature because it boosted performance of our initial testing, even without words been previously stemmed or lemmatized (stop words were excluded). Counts of maximum of 100 unigrams were used.
	\item \textit{Type-token ratio}. We wanted to use that feature to measure the vocabulary richness in a token's context, but after we finished with performance evaluations we realized that there was a bug in implementation and that feature did not contribute to overal results. The feature should be calculated as the ratio of vocabulary size and total number of tokens of the text \citep{stamatatos-2009a}.
\end{itemize}

We implemented two variants of our final model, first with only a basic feature extractor $f_\mathrm{b}$ whether in second feature transformation $f_\mathrm{t}$ was used. For the first variant, where only a basic feature extractor $f_\mathrm{b}$ was used, we extracted features for each document separately. In the variant with both basic extractor $f_\mathrm{b}$ and transformation $f_\mathrm{t}$, all documents were concatenated to a single one from which basic and transformed feature vectors were obtained. In both approaches features were extracted from a sliding window which included a context of each token.

@@

scaling!!!!!!

@@

The final step in our approach was clustering of obtained feature vectors which represented stylistic contexts of tokens. The number of clusters was equal to the number of authors in tasks $a$ and $b$, but in task $c$ it was unknown. We tested various clustering algorithms available in \texttt{scikit-learn}\footnote{\url{http://scikit-learn.org/stable/}} Python toolkit in all tasks. For tasks $a$ and $b$ we propose hierarchical agglomerative clustering, and for task $c$ DBSCAN algorithm which estimates the number clusters automatically. The reason of choosing that algorithms is described below.

@@

1

baselines

repo link

Mention cofidences.

feature init test
